A Process for the Emulation of Comparative Oncology Trials with Real-world Evidence (ENCORE)

### 2024-12-15 manuscript/manuscript.qmd
minor formatting and added word count

[**Manuscript word count:**]{.underline} xxx words / 8,000 words

[**Abstract word count:**]{.underline} xxx words / 250 words

@@

[**Supplementary material:**]{.underline} Supplementary tables and figures

@@

**Fit-for-purpose data.** Real-world data fitness and emulation feasibility for a given shortlisted candidate trial will be assessed in multiple steps based on guidance of the oncology quality, characterization, and assessment of real-world data (Oncology QCARD) Initiative.[@rivera2022friends] The first step assesses if relevant variables like exposure/line of therapy, outcomes, and covariates are generally available, measured and operationalizable in routine-care. Since a vast majority of oncological RCTs in recent years have focused on selected, biomarker-defined populations, subtleties in measurement and operationalizability of specific biomarkers must be reflected to ensure a representative and large enough study population. For example, immunotherapies have significantly changed the cancer treatment landscape since the approval of the first PD-L1 inhibitor in 2015. With many trials that have followed thereafter, the operationalization of the expression of the PD-L1 biomarker in RCTs (e.g., as a percent staining, tumor proportion score or combined positive score) has also evolved since then and PD-L1 '*positivity'* may have different definitions across calendar years based on different cut-off values.

@@

**Descriptives and data exploration.** Critical aspects when emulating oncology trials are the choice and estimation of the appropriate estimand of interest.[@rufibach2018] Particularly when emulating pivotal trials of paradigm-changing treatments, multiple aspects need to be considered such as the contemporaneity of the (historical) control cohort, the adoption rate of the novel intervention in routine care, the magnitude of the clinical treatment benefit and the rate in which (particularly patients in the control arm) discontinue or cross-over to the interventional treatment, which could lastly bias emulated treatment effects towards the null. To that end, comprehensive data explorations will be performed as part of the protocol development to contextualize these parameters and (if reported) draw comparisons to the emulated trial. Example for such standard diagnostics are visualized in @fig-initiators.

@@

-   **Regulatory agreement**: agreement between RCT and emulated trial treatment effect with regards to directionality and statistical significance. This agreement metric mimics the question if RWE would have lead to the same regulatory conclusion as the trial, based on standard regulatory decision rules.

-   **Estimate agreement**: agreement that the estimated RWE treatment treatment effect is within the 95% CI of the RCT treatment effect estimate. Provided that for some emulations, the power of the RWE study may be larger than that of the original RCT, this could lead to situations where there is no regulatory agreement although the treatment effect estimates are highly overlapping but with the RCT estimate crossing the null (or vice versa in case the RCT has a larger power than the RWE emulation).

-   **SMD agreement**: quantification of the agreement between the emulated RWE and RCT treatment effect estimate. The SMD is calculated as

>>>>

[**Manuscript word count:**]{.underline} 4,131 words / 8,000 words

[**Abstract word count:**]{.underline} 248 words / 250 words

@@

[**Supplementary material:**]{.underline} Supplementary figures and material

@@

#### **Fit-for-purpose data**

Real-world data fitness and emulation feasibility for a given shortlisted candidate trial will be assessed in multiple steps based on guidance of the oncology quality, characterization, and assessment of real-world data (Oncology QCARD) Initiative.[@rivera2022friends] The first step assesses if relevant variables like exposure/line of therapy, outcomes, and covariates are generally available, measured and operationalizable in routine-care. Since a vast majority of oncological RCTs in recent years have focused on selected, biomarker-defined populations, subtleties in measurement and operationalizability of specific biomarkers must be reflected to ensure a representative and large enough study population. For example, immunotherapies have significantly changed the cancer treatment landscape since the approval of the first PD-L1 inhibitor in 2015. With many trials that have followed thereafter, the operationalization of the expression of the PD-L1 biomarker in RCTs (e.g., as a percent staining, tumor proportion score or combined positive score) has also evolved since then and PD-L1 '*positivity'* may have different definitions across calendar years based on different cut-off values.

@@

#### **Descriptives and data exploration**

Critical aspects when emulating oncology trials are the choice and estimation of the appropriate estimand of interest.[@rufibach2018] Particularly when emulating pivotal trials of paradigm-changing treatments, multiple aspects need to be considered such as the contemporaneity of the (historical) control cohort, the adoption rate of the novel intervention in routine care, the magnitude of the clinical treatment benefit and the rate in which (particularly patients in the control arm) discontinue or cross-over to the interventional treatment, which could lastly bias emulated treatment effects towards the null. To that end, comprehensive data explorations will be performed as part of the protocol development to contextualize these parameters and (if reported) draw comparisons to the emulated trial. Example for such standard diagnostics are visualized in @fig-initiators.

@@

**Regulatory agreement**: agreement between RCT and emulated trial treatment effect with regards to directionality and statistical significance. This agreement metric mimics the question if RWE would have lead to the same regulatory conclusion as the trial, based on standard regulatory decision rules.

**Estimate agreement**: agreement that the estimated RWE treatment treatment effect is within the 95% CI of the RCT treatment effect estimate. Provided that for some emulations, the power of the RWE study may be larger than that of the original RCT, this could lead to situations where there is no regulatory agreement although the treatment effect estimates are highly overlapping but with the RCT estimate crossing the null (or vice versa in case the RCT has a larger power than the RWE emulation).

**SMD agreement**: quantification of the agreement between the emulated RWE and RCT treatment effect estimate. The SMD is calculated as
